Manifold Clustering in the Setting of Generalized Random Dot Product Graphs

### 2023-05-12 sdss-2023/poster.Rmd
minor adjustments to poster

# Block Models as GRDPGs

It has been previously shown that the SBM, DCBM, and PABM are GRDPGs in which the communities lie on point masses, line segments, and subspaces, respectively.

@@

```{r mbconnectome-ase, fig.height = 3.75, fig.width = 7, out.width = '100%', cache = TRUE}
Xhat.l <- embedding(A.l, 2, 1)
ase.l <- ggplot() + 
  geom_point(aes(x = Xhat.l[, 1], y = Xhat.l[, 2],
                 colour = factor(z.l))) + 
  coord_fixed(.5) +
  theme_bw() + 
  theme(legend.position = 'none') +
  scale_colour_brewer(palette = 'Set1') + 
  labs(x = NULL, y = NULL, colour = NULL) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

Xhat.r <- embedding(A.r, 2, 1)
ase.r <- ggplot() + 
  geom_point(aes(x = Xhat.r[, 1], y = Xhat.r[, 2],
                 colour = factor(z.r))) + 
  coord_fixed(.5) +
  theme_bw() + 
  theme(legend.position = 'none') +
  scale_colour_brewer(palette = 'Set1') + 
  labs(x = NULL, y = NULL, colour = NULL) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

@@

mbconnectome.out.r <- manifold.clustering(Xhat.r, 
                                          length(unique(z.r)), 
                                          initialization = z.init.r,
                                          intercept = FALSE, 
                                          parallel = TRUE, 
                                          verbose = FALSE, 
                                          animate = FALSE, 
                                          curve.init = 'x')
fitted.r <- plot.estimated.curves(Xhat.r, mbconnectome.out.r) + 
  scale_colour_brewer(palette = 'Set1') + 
  coord_fixed(.5) +
  labs(x = NULL, y = NULL) + 
  ylim(min(Xhat.r[, 2]), max(Xhat.r[, 2])) + 
  xlim(min(Xhat.r[, 1]), max(Xhat.r[, 1]))

mbconnectome.out.l <- manifold.clustering(Xhat.l, 
                                          length(unique(z.l)), 
                                          initialization = z.init.l,
                                          intercept = FALSE, 
                                          parallel = TRUE, 
                                          verbose = FALSE, 
                                          animate = FALSE, 
                                          curve.init = 'x')
fitted.l <- plot.estimated.curves(Xhat.l, mbconnectome.out.l) + 
  scale_colour_brewer(palette = 'Set1') + 
  coord_fixed(.5) +
  labs(x = NULL, y = NULL) + 
  ylim(min(Xhat.l[, 2]), max(Xhat.l[, 2])) + 
  xlim(min(Xhat.l[, 1]), max(Xhat.l[, 1]))

@@

Block models can be expressed as GRDPGs in which the communities are linear structures in the latent space. 
We propose the manifold block model to characterize nonlinear latent structures and the $K$-curves clustering algorithm to estimate these structures for community detection.

>>>>

# Block Models as Linear GRDPGs

It has been previously shown that the SBM, DCBM, and PABM are GRDPGs in which the communities correspond to linear structures in the latent space.

@@

```{r mbconnectome-ase, fig.height = 3.75, fig.width = 7, out.width = '100%', cache = TRUE}
Xhat.l <- embedding(A.l, 2, 1)
ase.l <- ggplot() + 
  geom_point(aes(x = Xhat.l[, 1], y = Xhat.l[, 2],
                 colour = factor(z.l))) + 
  # coord_fixed(.5) +
  theme_bw() + 
  theme(legend.position = 'none') +
  scale_colour_brewer(palette = 'Set1') + 
  labs(x = NULL, y = NULL, colour = NULL) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

Xhat.r <- embedding(A.r, 2, 1)
ase.r <- ggplot() + 
  geom_point(aes(x = Xhat.r[, 1], y = Xhat.r[, 2],
                 colour = factor(z.r))) + 
  # coord_fixed(.5) +
  theme_bw() + 
  theme(legend.position = 'none') +
  scale_colour_brewer(palette = 'Set1') + 
  labs(x = NULL, y = NULL, colour = NULL) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

@@

mbconnectome.out.r <- manifold.clustering(Xhat.r, 
                                          length(unique(z.r)), 
                                          initialization = z.init.r,
                                          intercept = FALSE, 
                                          parallel = TRUE, 
                                          verbose = FALSE, 
                                          animate = FALSE, 
                                          curve.init = 'x')
fitted.r <- plot.estimated.curves(Xhat.r, mbconnectome.out.r) + 
  scale_colour_brewer(palette = 'Set1') + 
  coord_fixed(.4) +
  labs(x = NULL, y = NULL) + 
  ylim(min(Xhat.r[, 2]), max(Xhat.r[, 2])) + 
  xlim(min(Xhat.r[, 1]), max(Xhat.r[, 1])) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

mbconnectome.out.l <- manifold.clustering(Xhat.l, 
                                          length(unique(z.l)), 
                                          initialization = z.init.l,
                                          intercept = FALSE, 
                                          parallel = TRUE, 
                                          verbose = FALSE, 
                                          animate = FALSE, 
                                          curve.init = 'x')
fitted.l <- plot.estimated.curves(Xhat.l, mbconnectome.out.l) + 
  scale_colour_brewer(palette = 'Set1') + 
  coord_fixed(.4) +
  labs(x = NULL, y = NULL) + 
  ylim(min(Xhat.l[, 2]), max(Xhat.l[, 2])) + 
  xlim(min(Xhat.l[, 1]), max(Xhat.l[, 1])) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

@@

Block models can be expressed as GRDPGs in which the communities are linear structures in the latent space. 
We propose the manifold block model to extend this to nonlinear latent structures and the $K$-curves clustering algorithm to estimate these structures for community detection.
